Re-finding to support a Deep Web Question Answering System

### 2010-03-05 trunk/paper/semsearch10/content/nlp.tex
More New stuff from jnw

% Joir-dan stuff

When we have completed our Wikipedia corpora from all the blankets in the category hierarchy, our system needs to make sense of the data for execution. The learning process involves the calculation of two n-gram frequency distributions: an n-gram/vocabulary distribution, and an n-gram/category distribution. The first value is the count of individual n-grams in the combined vocabulary of all corpora, and the second one is just the count of individual n-grams within a category’s blanket. The categories’ corpora reflect one long context per category, instead of breaking up sentences into different contexts. The idea behind this is that Wikipedia pages are already grouped into categories, that we can assume are related to the topics of each sentence within the corpus. We store distributions, rather than priors, primarily because if at anytime a blanket vocabulary seems inadequate, we can append terms from other sources to both distribution types and recalculate our values. We can simply calculate probabilities and store these separately for the execution of our natural query engine.

The goal, is to choose that category for use in the SSQ structure. This process requires the tagging and parsing of our natural query, the extraction of key elements, primarily adjectives and nouns, and then using n-gram priors to calculate the appropriate category associated with that term.

Picking out these terms requires knowledge of the structure of the question being passed to the system. The ideal situation would be that we could recognize question patterns, and decide where key information is within the question based on a given pattern. Currently, we only work with questions in an “answer” form, which means there is little to no transformation of the structure, to go from a question to an answer.

Example

Question: A 1997 Toyota Corolla needs what size tires?
Answer:   A 1997 Toyota Corolla needs 17” tires.

>>>>

\section{Natuarl Language Queries}
% Joir-dan stuff
The first step in answering a user query is to determine what, in
fact, is being asked.  Any system that supports user queries places
some constraints on the required input.  These constraints can be
relatively minimal (``Users must type their input on a keyboard in
French.'') or quite constrained (``Users must constuct fully
categorized semi-structured queries employing only those terms and
categories supplied by our database.'')  Our initial goal in Morpheus
is two-fold:

\begin{enumerate}
\item Implement an initial natural language interface that allows
  users to make queries in a loosely constrained English language
  subset.
\item Provide a framework that can support a robust
  interface that employs statistical language processing
  methods.  
\end{enumerate}

At present, we can process English language questions posed in answer
form.  This requires minimal structural transformation in order to
represent the semantic information contained within the question.  We
translate these questions into semi-structured queries that associate
English language terms and modifier phrases with contexts (\emph{who,
  what, when, where})

Input contexts of such a semi-structured query are generally given as
terms (particular dates, names of cars, etc.) while output contexts
are given as categories (tire type, tire size). Thus the inputs of the
user SSQ are uncategorized, while their outputs are categorized.  The
SSQs associated with a QRM are fully categorized (as will be discussed
in section \ref{sec:qrr}). To match a user SSQ to existing QRMs, we
must be able to match their SSQ input and output contexts.  This
requires matching the input terms to their categories, a process we
discuss in section \ref{sec:terms}.

### 2010-03-05 trunk/paper/semsearch10/content/qrr.tex
More New stuff from jnw

\section{Query Resolution Recorder (QRR)}
\label{sec:qrr}
% Chris
Answering a question using the deep web requires one to navigate
through a sequence of one or more web pages and links.  Many of the
accesses involve clicks through web forms to resulting pages.  We have
developed a model that represents the various actions a user may carry
out during this process.

During the answer retrieval process, a user discovers an element of
the query's result on a web page.  Next, she highlights that element,
typically a text string.  This highlight allows the QRR to record the
source of the query's answer.  The QRR stores all information from the
user query in our data store.  When the user submits a web form she
may associate each of the form's inputs with one of our stored context
classes.

The discovery process can be broken down into three cases:
\begin{enumerate}
\item The user typed in a URL and highlighted part of or all the
  answer.
\item The user typed in a URL, clicked a link, and found part or all
  of the answer.
\item The user typed in a URL, filled out a form, and found part or
  all of the answer on a subsequent page.
\end{enumerate}

Any discovery process consists of a sequence of these transactions.
Additionally, data collected for inputs and outputs may be found
during any stage of the complete discovery process.  An output
collected during one stage of an extraction may be the input to a
subsequent stage in the extraction.  The QRR builds the completed user
collection process into a QRM.  The following is a mathematical
formulation of information collected for user web interactions.

A QRM is a 5-tuple $Z = \left< \Upsilon, \Omega, P, M, R, A \right>$ such that:

\begin{enumerate}

\item $\Upsilon$ is the sequence of input categories $\left<
  \Upsilon_{1}, \Upsilon_{2}, \Upsilon_{3}, \Upsilon_{4} \right>$
  where $\Upsilon_{1}$ represents the \emph{who} context of the
  associated query, $\Upsilon_{2}$ represents the \emph{what} context,
  $\Upsilon_{3}$ represents the \emph{when} context, and
  $\Upsilon_{4}$ represents the \emph{where} context.

\item $\Omega$ is the sequence of output categories $\left<\Omega_{1},
  \Omega_{2}, \Omega_{3}, \Omega_{4}\right>$ where $\Omega_{1}$
  represents the \emph{who} context of the associated query,
  $\Omega_{2}$ represents the \emph{what} context, $\Omega_{3}$
  represents the \emph{when} context, and $\Omega_{4}$ represents the
  \emph{where} context.

@@

\item $R$ is an ontological realm.

@@

%\subsection{Generalization}
% Christan
The response pages created by querying web forms are dynamically
generated with a templated structure. This generalization per page is
done using the GenPath algorithm developed by Badica et
al. \cite{Badica06}. Additionally, the community may be able to answer
a particular SSQ using many different page interactions. Therefore, we
use a method of generalizing extraction path to a result across
pages. This method removes unnecessary page interactions and finds the
shortest path to the result pages.

Websites wrapped by a system such as this are highly susceptible to
structural changes \cite{TanZMG07}. Generalizing the extraction paths
helps overcome small structural changes. If a website changes to the
point where a QRM is no longer useful, the QRM is not considered
fresh. User ratings of QRM freshness can be employed in the SSQ matching
process detailed in Section \ref{sec:ssq}.

>>>>

\section{Query Resolution Recorder (QRR)}
\label{sec:qrr}
% Chris
Answering a question using the deep web requires one to navigate
through a sequence of one or more web pages. Many of the accesses
involve clicks through web forms or links to resulting pages.  We have
developed a model that represents the various actions a user may carry
out during this process.

Web interactions can be broken down into three types:
\begin{enumerate}
\item The user fills in and submits a web form.
\item The user clicks a link.
\item The user highlights text that is needed in the query or forms
  part of the answer.
\end{enumerate}

When a users submit a web form they are prompted for the classes of
each of the form inputs. Here, class refers to categorization of data,
e.g. name, date, etc. These classes allow the QRR to map the user's
input SSQ tokens (which are also categorized) to their usage in a
given form. For a clicked link, the QRR simply records the target url
and nothing more. Finally, during the answer retrieval process when a
user discovers all or part of the desired information on a web page,
she highlights the text in question. This highlight allows the QRR to
record the source of the query's answer. All information gathered
during the recording process is saved in our data store.

Any query answering process consists of a sequence of these transactions.
Additionally, data collected for inputs and outputs may be found
during any stage of the complete query process.  An output
collected during one stage of an extraction may be the input to a
subsequent stage in the extraction.  The QRR builds the completed user
collection process into a QRM.

Conceptually, a QRM is the data structure that models the question
answering process.  Each QRM contains an ontilogical realm, an SSQ,
and information to support the query answering process.  The realm may
be one or more descriptive categories and is used for classification
of the QRMs.  The SSQ describes the set of possible input and output
categories from the associated natural language query.  The possible
set of phrase inputs and outputs are grouped into contextual bins
labeled \emph{who}, \emph{what}, \emph{when} and \emph{where}.  The
ontilogical realm and the SSQ parameters act as a signature for the
QRM object, analogous to a function signature.  The ontological
representation of the SSQ is described in section \ref{sec:ssq}.

Additionally, the QRM contains representations for every page visited
and the order of visitation.  For each page, we store the list of
inputs to the URL such as form inputs and the possible outputs
(highlighted portions of the webpage).  There is a mapping between the
SSQ inputs and outputs used on any page to the specific page input.
Finally, there is a structure with mappings to the user inputs or
outputs that are displayed to the user as answers.

The following is a mathematical description of the information
collected for user web interactions.

A QRM is a 4-tuple $Z = \left< \Sigma, P, M, A
\right>$ with the following structure:

\begin{enumerate}
%%%%%%%%%%%%%%% Separate the first 2 points as descriptions of the SSQ. --------------

\item $\Sigma = ( R, \Upsilon, \Omega )$ is the SSQ associated with the QRM.

  $R$ is the ontological realm associated with the QRM.

  $\Upsilon$ is the sequence of input categories drawn from $R$
  $\left< \Upsilon_{1}, \Upsilon_{2}, \Upsilon_{3}, \Upsilon_{4}
  \right>$ where $\Upsilon_{1}$ represents the \emph{who} context of
  the associated query, $\Upsilon_{2}$ represents the \emph{what}
  context, $\Upsilon_{3}$ represents the \emph{when} context, and
  $\Upsilon_{4}$ represents the \emph{where} context.

  $\Omega$ is the sequence of output categories drawn from $R$
  $\left<\Omega_{1}, \Omega_{2}, \Omega_{3}, \Omega_{4}\right>$ where
  $\Omega_{1}$ represents the \emph{who} context of the associated
  query, $\Omega_{2}$ represents the \emph{what} context, $\Omega_{3}$
  represents the \emph{when} context, and $\Omega_{4}$ represents the
  \emph{where} context.

@@

@@

%\subsection{Generalization}
% Christan

The response pages created by querying web forms are dynamically
generated with a template structure. This template structure contains
XPath [!!!!!need reference] selectors to access particular web page
elements.  One could employ XPath selectors that exactly match the
selection paths to the web page elements accessed by the guide during
query recording, however, using such specific paths fails to provide
robust performance because web sites are highly susceptible to
structural changes \cite{TanZMG07}.  To overcome this difficulty, we
generalize these XPaths using the \emph{GenPath algorithm} developed by
Badica et al. \cite{Badica06}.

Generalizing the extraction paths helps overcome small structural
changes. However, If a website changes to the point where a QRM is no
longer useful, the QRM is not considered fresh. User ratings of QRM
freshness can be employed in the SSQ matching process detailed in
Section \ref{sec:oqom}.

In many cases, the same information may be found by a number of guides
from the same deep web source, but by a different navigation path.
Therefore, we also employ a method of generalizing extraction paths
across page interactions. By doing this, one can remove unnecessary
page interactions and finds the most effective navigation path
necessary to produce a result.
%-------------------------------------------------------------------------------

### 2010-03-05 trunk/paper/semsearch10/content/results_futurework_conclusion.tex
More New stuff from jnw

\section{Future Work}
- Talk about the offline execution mode and class structural matching algorithm
- Researchers have stated 39, 57, and 63 percent of web queries are unique \cite{1277770,331405,621942}.  We believe that an even higher number of those are parameterizable, meaning, a set of queries are all related and may be answered using a single QRM.  We plan to investigate this.

@@

>>>>

\section{Future Work}

The greatest limitation to our current work is the diversity of the
QRR.  In order to efficiently answer new user questions the systems
needs a large amount of user annotated queries.  However, a
significant number of user interactions is required to understand the
answer retrieval process.  By continuing to study the methods of user
data collection we plan to develop a model to support an unsupervised
question resolution process.  This will increase the scalability and
efficiency of the system.

Because each QRM may be treated as a function, we plan on leveraging
previous work to allow new questions to be solved by connecting
together several QRMs \cite{morpheus1, transformscout}.  This method
would expand the types of queries that are answerable by the system.

Currently, the category heterarchy generated from the DBpedia
categories and our SSQ ontologies is stored in separate XML files. We
are in the process of moving these ontologies into an efficient
ontology database. Similarly, employing user input brings about the
likelihood that redundant SSQ ontological elements will be inserted
into our ontology store. We are working on merging the similar SSQs
[\ref{sec:ssq}] based on the SSQ realms and other ontological
structural information so that we can remove the redundancy in the
database.

We cannot rely on the DBpedia categories and the associated Wikipedia
pages to provide corpora for all our realms.  The categories in
DBpedia store and documents in the Wikipedia are insufficient to
characterize the linguistic space of queries belonging to many
different realms.  Reasonable prior probabilities for the names of
musical groups within the category \emph{musical groups}, for example,
will be significantly different from what would be manufactured by
simply finding n-gram frequency within the related Wikipedia
documents. We are investigating methods to identify categories within
the deep web and to find new sources to construct \textit{weak} realms
in our ontology hierarchy.

%% I may have misstated the statistics that are unique/repeats
% need to clarify the statistics 
%% Add language about how we 
Researchers have stated that somewhere from 39 to 63 percent of web
queries are unique \cite{1277770,331405,621942}.  We believe that a
large percentage of the stated unique queries are in-fact related and
parameterizable, meaning, several subsets of unique queries may be all
related and may be answered using a single QRM.  It is difficult to
discover this property between queries.  Previous work suggest
techniques to find related queries but it will take new efficient
methods to discover parameterizable queries.

@@

In this paper we provide a discussion on a new type of question
answering system that uses previously answered user questions.  We
described our method of modeling a users question answering process
and how we are able to discover semantically similar queries.

### 2010-03-05 trunk/paper/semsearch10/content/ssq_ranking.tex
More New stuff from jnw

\section{SSQ Ranking}
\label{sec:ssq}
In order to match a new query with a SSQ within our data store, we
have defined similarity measures between the ontologies of the new
query and the candidate SSQs. These measures capture the distance
between the ontology classes, properties, and individuals at the
conceptual level in addition to simply matching URIs. The similarity
measures are implemented in two scenarios: \textit{on-line execution}
and \textit{off-line mode}. The \textit{on-line execution} is time
sensitive and is used primarily for relating newly defined SSQs to the
SSQs in the store. During \textit{off-line execution}, the system
clusters similar SSQs.

The similarity measure for the SSQ ontologies is determined by the
following components.
\begin{enumerate}
    \item QueryClass: an OWL class and the root of the SSQ ontology
      structure.
    \item hasRealm: an OWL object property with its domain as the SSQ
      Query class and its range as a realm class.
    \item Contexts such as who, what, when, and where are OWL object
      properties, whose domains are the query classes and ranges are
      OWL classes.
    \item SSQClass: an OWL class which is created from the components
      described above.
\end{enumerate}
Based on this ontology structure and the execution scenarios, we have
created an algorithm for finding the matches between SSQ
ontologies. It is based on class divergence, which is motivated by the
concept of multiple dispatch in CLOS \cite{95411} and Dylan
programming \cite{DBLP:conf/oopsla/BarrettCHMPW96} for generic
function type matches. We consider the realm class and context
property ranges for determining SSQ similarity. Once we have the similarity measures between the user defined SSQ and relevant QRMs in the store, we can order the relevant QRMs. The order provides a ranking for the results to the user. Subsections \ref{sec:ctd}, and \ref{sec:oqom} describe about this in detail.

@@

\begin{equation}
d_{class}(S, T) = \begin{cases}
0 & if S.{uri} \equiv T.{uri}\\
d(S, T) & if S \prec T\\
1 & if T \prec S\\
d(S,root) \\+ d(S,C) + d(T,C) & otherwise
\end{cases}
\end{equation}

We normalize $d_{class}(S, T)$ by dividing by three times the height
of the ontology tree.

@@

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=90mm]{class_divergence.eps}
\caption{Abbrv. graphical representation of an automotive ontology}
\label{fig:class_divergence}
\end{figure}

Figure \ref{fig:class_divergence} illustrates the calculation of class
divergence. This sample abbreviated ontology that was created from the
DBpedia category heirarchy is rooted at the class
\textit{DBpediaRoot}. The arrows represent the inheritance
relationship between the classes in the topology.

Suppose we want to find the divergence $d_{class}$\textit{(Automotive
  Industry, Racing Cars)}. The \textit{Automotive Industry} is the
subclass of \textit{Automobiles} and \textit{Racing Cars} is the
subclass of \textit{Auto Racinng}. Therefore, the common ancestor for
these two can be determined as \textit{Automobiles} ($C$). They share
the path from \textit{Automobiles} to the $root$. So, the unnormalized
divergence \textit{d(Automotive Industry, root)} is 32 (calcuted from
the path to root), \textit{d(Automotive Industry, Automobiles)} is one
(dashed arrow), and \textit{d(Racing Cars, Automobiles)} is two
(dashed arrow). Suppose, the ontology tree height is calculated as
294. Thus, \textit{d(Automotive Industry, Racing Cars)} is 35, which
corresponds to a normalized value of $35/294$.

@@

We use class divergence to calculate the relevance between a source
SSQ and a target SSQ associated with a QRM.  Each SSQ will have four
input classes, four output classes, and one realm.  We refer to the
sequence of these classes $\langle \Upsilon_1, ..., \Upsilon_4,
\Omega_1, ..., \Omega_4, R \rangle$ as the signature of an SSQ. If an
input or output class is unspecified in the query, it is associated
with the class $bottom$, which has no subclasses or individuals.

The relevance of QRM $Q$ with signature $\langle \Upsilon_{Q_1}, ...,
\Upsilon_{Q_4}, \\\Omega_{Q_1}, ... , \Omega_{Q_4}, R_Q\rangle$ to SSQ
$S$ with signature $\langle \Upsilon_{S_1}, ..., \Upsilon_{S_4},
\\\Omega_{S_1}, ... , \Omega_{S_4}, R_S \rangle$ is

\begin{eqnarray}
\label{qrmsimilarity}
 R(S,Q) = 1 - (d_{R} + d_{\Upsilon} + d_{\Omega})
\end{eqnarray}
\noindent where 
\begin{eqnarray}
d_{R} = \pi_9 d_{class}(R_S,R_Q) \nonumber
\end{eqnarray}
\begin{eqnarray}
d_{\Upsilon}  = \sum_{k=1}^4 {\pi_k} {d_{class}(\Upsilon_{S_k},\Upsilon_{Q_k})} \nonumber
\end{eqnarray}
\begin{eqnarray}
 d_{\Omega} = \sum_{j=1}^4 \pi_{4+j} d_{class}(\Omega_{S_j},\Omega_{Q_j})  \nonumber
\end{eqnarray}
\noindent the weights ${\pi}_k$ of the property hasRealm and other context properties satisfies the conditions 
\begin{equation}
\label{piconstraint1}
\sum_{k=1}^{9} {{\pi}_k} = 1
\end{equation}
\noindent and for $1 \leq i \leq 9$

\begin{equation}
\label{piconstraint2}
{\pi}_i > 0
\end{equation}

>>>>

\section{SSQ Ranking}
\label{sec:ssq}
In order to match a new query with a SSQ within our data store, we
have defined similarity measures between the ontologies of the new
query and the candidate SSQs. These measures capture the distance
between the ontology classes, properties, and individuals at the
conceptual level in addition to simply matching URIs. The similarity
measures are implemented in two scenarios: \textit{on-line execution}
and \textit{off-line mode}. The \textit{on-line execution} is time
sensitive and is used primarily for relating newly defined SSQs to the
SSQs in the store. During \textit{off-line execution}, the system
clusters similar SSQs.
% Move to QRR or make a reference  
The similarity measure for the SSQ ontologies is determined by the
following components.
\begin{enumerate}
    \item QueryClass: an OWL class and the root of the SSQ ontology
      structure.
    \item hasRealm: an OWL object property with its domain as the SSQ
      Query class and its range as a realm class.
    \item Contexts such as who, what, when, and where are OWL object
      properties, whose domains are the query classes and ranges are
      OWL classes.
    \item SSQClass: an OWL class which is created from the components
      described above.
\end{enumerate}
Based on this ontology structure and the execution scenarios, we have
created an algorithm for finding the matches between SSQ
ontologies. It is based on class divergence, which is motivated by the
concept of multiple dispatch in CLOS \cite{95411} and Dylan
programming \cite{DBLP:conf/oopsla/BarrettCHMPW96} for generic
function type matches. We consider the realm class and context
property ranges for determining SSQ similarity. Once we have the similarity measures between the user defined SSQ and relevant QRMs in the store, we can order the relevant QRMs. The order provides a ranking for the results to the user. Subsections \ref{sec:ctd}, and \ref{sec:oqom} describe about this in detail.

@@

\begin{equation}
d_{class}(S, T) = \begin{cases}
0 & if S.{Uri} \equiv T.{Uri}\\
d(S, T)/3h & if S \prec T\\
1 & if T \prec S\\
(d(S,root) + d(S,C) \\ \ \ \ \ + d(T,C))/3h & otherwise
\end{cases}
\end{equation}

where $h$ is three times the height of the ontology tree.

@@

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=90mm]{class_divergence.eps}
\caption{Abbrev. graphical representation of an automotive ontology}
\label{fig:class_divergence}
\end{figure}

Figure \ref{fig:class_divergence} illustrates the calculation of class
divergence. This sample abbreviated ontology that was created from the
DBpedia category hierarchy is rooted at the class
\textit{DBpediaRoot}. The arrows represent the inheritance
relationship between the classes in the topology.

Suppose we want to find the divergence $d_{class}$\textit{(Automotive
  Industry, Racing Cars)}. The \textit{Automotive Industry} is the
subclass of \textit{Automobiles} and \textit{Racing Cars} is the
subclass of \textit{Auto Racing}, which is a subclass of \textit{Automobiles}. Therefore, the common ancestor for
these two can be determined as \textit{Automobiles} ($C$). They share
the path from \textit{Automobiles} to the $root$. So, the unnormalized
divergence \textit{d(Automotive Industry, root)} is 32 that is calculated from the path to the root \textit{DBpediaRoot}, \textit{d(Automotive Industry, Automobiles)} is one
(dashed arrow), and \textit{d(Racing Cars, Automobiles)} is two
(dashed arrow). Suppose, the ontology tree height is calculated as
294. Thus, $d_{class}$\textit{(Automotive Industry, Racing Cars)} is 35, which corresponds to a normalized value of $35/294$.

@@

We use class divergence to calculate the relevance between a new SSQ
and a SSQ associated with a QRM in the store.  Each SSQ will have four
input classes, four output classes, and one realm.  We refer to the
sequence of these classes $\langle \Upsilon_1, ..., \Upsilon_4,
\Omega_1, ..., \Omega_4, R \rangle$ as the signature of an SSQ. If an
input or output class is unspecified in the query, it is associated
with the class $\perp$, which is a subclass of every other class has
no subclasses or individuals.

The relevance of QRM $Q$ with signature $\langle \Upsilon_{Q_1}, ...,
\Upsilon_{Q_4}, \\\Omega_{Q_1}, ... , \Omega_{Q_4}, R_Q\rangle$ to SSQ
$S$ with signature $\langle \Upsilon_{S_1}, ..., \Upsilon_{S_4},
\\\Omega_{S_1}, ... , \Omega_{S_4}, R_S \rangle$ is
\begin{eqnarray}
\label{qrmsimilarity}
 R(S,Q) = 1 - (d_{R} + d_{\Upsilon} + d_{\Omega})
\end{eqnarray}
\noindent where 
\begin{eqnarray}
d_{R} = \pi_9 d_{class}(R_S,R_Q) 
\end{eqnarray}
\begin{eqnarray}
d_{\Upsilon}  = \sum_{k=1}^4 {\pi_k} {d_{class}(\Upsilon_{S_k},\Upsilon_{Q_k})} 
\end{eqnarray}
\begin{eqnarray}
 d_{\Omega} = \sum_{j=1}^4 \pi_{4+j} d_{class}(\Omega_{S_j},\Omega_{Q_j}) 
\end{eqnarray}
\noindent the weights ${\pi}_i$ of the property hasRealm and other context properties satisfies the conditions
\begin{equation}
\label{piconstraint1}
\sum_{i=1}^{9} {{\pi}_i} = 1
\end{equation}
\begin{equation}
\label{piconstraint2}
{\pi}_i > 0
\end{equation}
\noindent and for $1 \leq i \leq 9$
